Reinforcement Learning Produces Dominant Strategies for the Iterated Prisoner's Dilemma

### 2017-08-10 plos_latex_template.tex
Write plos style version of the paper.
Restructure of file system is still needed:

- Copy tex tables in to document;
- Copy appendix in to document;
- Move images to root level.

\documentclass{article}

% Packages
\usepackage[margin=1.5cm, includefoot, footskip=30pt]{geometry}
\usepackage{hyperref}
\usepackage{multicol}
\usepackage{booktabs}
\usepackage{xstring}
\usepackage{subcaption}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{standalone}
\usepackage{fancyvrb}
\usepackage{authblk}

\usepackage{pdflscape}

\usepackage{tikz}
\usetikzlibrary{calc, shapes, patterns}

% Title
\title{Reinforcement Learning Produces Dominant Strategies for the
Iterated Prisoner's Dilemma}
\author[1]{Marc Harper}
\author[2]{Vincent Knight}
\author[3]{Martin Jones}
\author[4]{Georgios Koutsovoulos}
\author[2]{Nikoleta E. Glynatsi}
\author[3]{Owen Campbell}

\affil[1]{Google Inc., Mountain View, CA, USA}
\affil[2]{Cardiff University, School of Mathematics, UK}
\affil[3]{Not affiliated}
\affil[4]{INRA, Université Côte d‘Azur, CNRS, ISA, France}
\date{}

\begin{document}

\maketitle

\begin{abstract}
    We present tournament results and several powerful strategies for the Iterated
    Prisoner's Dilemma created using reinforcement learning techniques
    (evolutionary and particle swarm algorithms). These strategies are
    trained to perform well against a corpus of over 170 distinct
    opponents, including many well-known and classic strategies. All
    the trained strategies win standard tournaments against the total collection
    of other opponents. The trained strategies and one particular human made
    designed strategy are the top performers in noisy tournaments also.
\end{abstract}

\section{Introduction}\label{sec:introduction}

@@

Since Robert Axelrod's seminal tournament \cite{Axelrod1980a}, a number of
IPD\@ tournaments have been undertaken and are summarised in
Table~\ref{tbl:tournaments}. Further to the work described in
\cite{knight2016open} a regular set of standard, noisy~\cite{Bendor1991} and
probabilistic ending~\cite{Axelrod1980b} tournaments are carried out as more
strategies are added to the Axelrod library.
Details and results are available here:
\url{http://axelrod-tournament.readthedocs.io}. This work presents a detailed
analysis of a tournament with 176 strategies (details given
in Section~\ref{sec:results}).

\begin{table}[!hbtp]
    \begin{center}
        \begin{tabular}{ccccc}
            \toprule
            Year     & Reference                  & Number of Strategies & Type     & Source Code\\
            \midrule
            1979     & \cite{Axelrod1980a}        & 13                   & Standard & Not immediately available\\
            1979     & \cite{Axelrod1980b}        & 64                   & Standard & Available in FORTRAN\\
            1991     & \cite{Bendor1991}          & 13                   & Noisy    & Not immediately available\\
            2002     & \cite{Stephens2002}        & 16                   & Wildlife & Not applicable\\
            2005     & \cite{kendall2007iterated} & 223                  & Varied   & Not available \\
            2012     & \cite{Stewart2012}         & 13                   & Standard & Not fully available \\
            2016     & \cite{knight2016open}       & 129                  & Standard & Fully available \\
            \bottomrule
        \end{tabular}
    \end{center}
    \caption{An overview of a selection of published tournaments. Not all
             tournaments were `standard' round robins; for more details
             see the indicated references.}\label{tbl:tournaments}
\end{table}

@@

\section{The Strategy Archetypes}

The Axelrod library now contains many parametrised strategies trained using
machine learning
methods. Most are deterministic, use many rounds of memory, and perform
extremely well in tournaments as will be discussed in Section~\ref{sec:results}.
Training of these strategies will be discussed in Section~\ref{sec:methods}.
These strategies can encode a variety
of other strategies, including classic strategies like Tit For Tat
\cite{Axelrod1980},
handshake strategies, and grudging strategies, that always defect after
an opponent defection.

\subsection{LookerUp}\label{sec:lookerup}

The LookerUp strategy is based on a lookup table and encodes a set of
deterministic responses based on the opponent's first $n_1$ moves, the
opponent's last $m_1$ moves, and the players last $m_2$ moves. If $n_1 > 0$ then
the player has infinite memory depth, otherwise it has depth $\max(m_1, m_2)$.
This is illustrated diagrammatically in Figure~\ref{fig:lookerup}.

\begin{figure}[!hbtp]
    \centering
    \includestandalone[height=.3\textheight]{./assets/lookerup}
    \caption{Diagrammatic representation of the Looker up Archetype.}
    \label{fig:lookerup}
\end{figure}

@@

\subsection{Gambler}\label{sec:gambler}

Gambler is a stochastic variant of LookerUp. Instead of deterministically
encoded moves the lookup table emits probabilities which are
used to choose cooperation or defection.
This is illustrated diagrammatically in Figure~\ref{fig:gambler}.

\begin{figure}[!hbtp]
    \centering
    \includestandalone[height=.3\textheight]{./assets/gambler}
    \caption{Diagrammatic representation of the Gambler Archetype.}
    \label{fig:gambler}
\end{figure}

@@

\subsection{ANN: Single Hidden Layer Artificial Neural Network}\label{sec:ann}

@@

\begin{multicols}{2}
    \begin{itemize}
        \item Opponent's first move is C
        \item Opponent's first move is D
        \item Opponent's second move is C
        \item Opponent's second move is D
        \item Player's previous move is C
        \item Player's previous move is D
        \item Player's second previous move is C
        \item Player's second previous move is D
        \item Opponent's previous move is C
        \item Opponent's previous move is D
        \item Opponent's second previous move is C
        \item Opponent's second previous move is D
        \item Total opponent cooperations
        \item Total opponent defections
        \item Total player cooperations
        \item Total player defections
        \item Round number
    \end{itemize}
\end{multicols}

These are then input into a feed forward neural network with one layer and
user-supplied width.  This is illustrated diagrammatically in
Figure~\ref{fig:ann}.

\begin{figure}[!hbtp]
    \centering
    \includestandalone[height=.3\textheight]{./assets/ann}
    \caption{Diagrammatic representation of the ANN Archetype.}
    \label{fig:ann}
\end{figure}

@@

\subsection{Finite State Machines}\label{sec:fsm}

Strategies based on finite state machines are deterministic and computationally efficient.
In each round of play the strategy selects an action based on the current state
and the opponent's last action, transitioning to a new state for the next round.
This is illustrated diagrammatically in Figure~\ref{fig:fsm}.

\begin{figure}[!hbtp]
    \centering
    \includestandalone[height=.3\textheight]{./assets/fsm}
    \caption{Diagrammatic representation of the Finite State Machine Archetype.}
    \label{fig:fsm}
\end{figure}

Training this strategy corresponds to finding mappings of states and histories
to an action and a state. Figure~\ref{fig:fsm_images} shows two of the trained
finite state machines. The layout of state nodes is kept the same between
Figure~\ref{fig:fsm16} and~\ref{fig:fsm16noise} to highlight the effect of
different training environments. Note also that two of the 16 states are not
used, this is also an outcome of the training process.

\begin{figure}[!hbtp]
    \centering
    \begin{subfigure}[t]{.5\textwidth}
        \includegraphics[height=.3\textheight]{./assets/FSM16.pdf}
        \caption{Evolved\_FSM\_16: trained to maximise score in a standard
        tournament}
        \label{fig:fsm16}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{.5\textwidth}
        \centering
        \includegraphics[height=.35\textheight]{./assets/FSM16noise.pdf}
        \caption{Evolved\_FSM\_16\_Noise\_05: trained to maximise score in a
        noisy tournament}
        \label{fig:fsm16noise}
    \end{subfigure}%
    \caption{Trained sixteen state Finite State Machine players.}
    \label{fig:fsm_images}
\end{figure}

\subsection{Hidden Markov Models}\label{sec:hmm}

A variant of finite state machine strategies are called hidden Markov models
(HMMs). Like the strategies based on finite state machines, these strategies
also encode an internal state. However, they use probabilistic transitions based on the
prior round of play to other states and cooperate or defect with various
probabilities at each state. This is
shown diagrammatically in Figure~\ref{fig:hmm}. Training this strategy
corresponds to finding mappings of states and histories to probabilities of
cooperating as well as probabilities of the next internal state.

\begin{figure}[!hbtp]
    \centering
    \includestandalone[height=.3\textheight]{./assets/hmm}
    \caption{Diagrammatic representation of the Hidden Markov Model Archetype.}
    \label{fig:hmm}
\end{figure}

\subsection{Meta Strategies}

@@

\section{Results}\label{sec:results}

@@

A total of \input{./assets/number_of_players.tex}are included, of which
\input{./assets/number_of_stochastic_players.tex}are stochastic.
Section~\ref{sec:standard} is concerned with the standard tournament with 200
turns whereas in Section~\ref{sec:noise} a tournament with 5\% noise is
discussed. Due to the inherent stochasticity of these IPD tournaments, these
tournaments were repeated \input{./assets/standard_number_of_repetitions.tex}
times. This allows for a detailed and confident analysis of the performance of
strategies. To illustrate the results considered,
Figure~\ref{fig:tit_for_tat_scores} shows the distribution of the mean score per
turn of Tit For Tat over all the repetitions. Similarly,
Figure~\ref{fig:tit_for_tat_ranks} shows the ranks of of Tit For Tat for each
repetition. (We note that it never wins a tournament). Finally
Figure~\ref{fig:tit_for_tat_wins} shows the number of opponents beaten in any given
tournament: Tit For Tat does not win any match. (This is due to the fact that it
will either draw with mutual cooperation or defect second).

\begin{figure}[!hbtp]
    \centering
    \begin{subfigure}[t]{.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./assets/standard_tft_scores.pdf}
        \caption{Scores}
        \label{fig:tit_for_tat_scores}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./assets/standard_tft_ranks.pdf}
        \caption{Ranks}
        \label{fig:tit_for_tat_ranks}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./assets/standard_tft_wins.pdf}
        \caption{Wins}
        \label{fig:tit_for_tat_wins}
    \end{subfigure}%

    \caption{Results for Tit For Tat over
        \protect\input{./assets/standard_number_of_repetitions.tex}tournaments.}
\end{figure}

@@

\subsection{Standard Tournament}\label{sec:standard}

@@

\begin{table}[!hbtp]
        \centering
        \input{./assets/standard_top_15_scores_summary.tex}
        \caption{Standard Tournament: Mean score per turn of top 15 strategies
            (ranked by median over
        \protect\input{./assets/standard_number_of_repetitions.tex}tournaments).
        The leaderboard is dominated by the trained strategies (indicated by a
        $^{*}$).}
        \label{tbl:standard_score}
\end{table}

For completeness, violin plots showing the distribution of the scores of each
strategy (again ranked by median score) are shown in
Figure~\ref{fig:standard_boxplot}.

\begin{landscape}
    \begin{figure}[!hbtp]
        \centering
        \includegraphics[width=\paperwidth]{./assets/standard_scores_boxplots.pdf}
        \caption{Standard Tournament: Mean score per turn (ranked by median
        over
        \protect\input{./assets/standard_number_of_repetitions.tex}tournaments).}
        \label{fig:standard_boxplot}
    \end{figure}
\end{landscape}

Pairwise payoff results are given as a heatmap (Figure~\ref{fig:standard_heatmap})
which shows that many strategies achieve mutual cooperation (obtaining a score
of 3). The top performing
strategies never defect first yet are able to exploit weaker strategies that
attempt to defect.

\begin{figure}[!hbtp]
    \centering
    \includegraphics[width=\textwidth]{./assets/standard_scores_heatmap.pdf}
    \caption{Standard Tournament: Mean score per turn of row players against
    column players (ranked by median over
        \protect\input{./assets/standard_number_of_repetitions.tex}tournaments).}
    \label{fig:standard_heatmap}
\end{figure}

The strategies that win the most matches
(Table~\ref{tbl:standard_wins_top_winners}) are Defector~\cite{Axelrod1984} and Aggravater~\cite{axelrodproject}, followed
by handshaking and zero determinant strategies~\cite{Press2012}.
This includes two handshaking
strategies that were the result of training to maximize Moran process fixation
(TF1 and TF2). No strategies were trained specifically to win matches. None of
the top scoring strategies appear in the top 15 list of strategies ranked by
match wins. This can be seen in Figure~\ref{fig:standard_winplot} where the
distribution of the number of wins of each strategy is shown.

\begin{table}[!hbtp]
    \centering
        \input{./assets/standard_top_15_winners_wins_summary.tex}
        \caption{Standard Tournament: Number of wins per tournament
        of top 15 strategies (ranked by median wins over
        \protect\input{./assets/standard_number_of_repetitions.tex}tournaments).}
        \label{tbl:standard_wins_top_winners}
\end{table}

\begin{landscape}
    \begin{figure}[!hbtp]
        \centering
        \includegraphics[width=\paperwidth]{./assets/standard_wins_boxplots.pdf}
        \caption{Standard Tournament: number of wins per tournament (ranked by
        median over
        \protect\input{./assets/standard_number_of_repetitions.tex}tournaments).}
        \label{fig:standard_winplot}
    \end{figure}
\end{landscape}

@@

\begin{table}[!hbtp]
    \centering
        \input{./assets/standard_top_15_wins_summary.tex}
        \caption{Standard Tournament: Number of wins per tournament
        of top 15 strategies (ranked by median score over
        \protect\input{./assets/standard_number_of_repetitions.tex}tournaments).}
        \label{tbl:standard_wins}
\end{table}

Finally, Table~\ref{tbl:standard_ranks} and
Figure~\ref{fig:standard_ranks_boxplot} show the ranks (based on median score)
of each strategy over the repeated tournaments. Whilst there is some
stochasticity, the top three strategies almost always rank in the top three. For
example, the worst that the Evolved Lookerup 2 2 2 ranks in any tournament
is 8th.

\begin{table}[!hbtp]
    \centering
        \input{./assets/standard_top_15_ranks_summary.tex}
        \caption{Standard Tournament: Rank in each tournament
        of top 15 strategies (ranked by median over
        \protect\input{./assets/standard_number_of_repetitions.tex}tournaments).}
        \label{tbl:standard_ranks}
\end{table}

>>>>

% Template for PLoS
% Version 3.4 January 2017
%
% % % % % % % % % % % % % % % % % % % % % %
%
% -- IMPORTANT NOTE
%
% This template contains comments intended
% to minimize problems and delays during our production
% process. Please follow the template instructions
% whenever possible.
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% Once your paper is accepted for publication,
% PLEASE REMOVE ALL TRACKED CHANGES in this file
% and leave only the final text of your manuscript.
% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
%
%
% There are no restrictions on package use within the LaTeX files except that
% no packages listed in the template may be deleted.
%
% Please do not include colors or graphics in the text.
%
% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
% TODO Move copy all the tables in to the file (JEEEEEEZ)
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% -- FIGURES AND TABLES
%
% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file.
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission.
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
% For figure citations, please use "Fig" instead of "Figure".
% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - spacing/line breaks within cells to alter layout or alignment
% - do not nest tabular environments (no tabular environments within tabular environments)
% - no graphics or colored text (cell background color/shading OK)
% See http://journals.plos.org/plosone/s/tables for table guidelines.
%
% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
%
% IMPORTANT
% Below are a few tips to help format your equations and other special characters according to our specifications. For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
%
% For inline equations, please be sure to include all portions of an equation in the math environment.  For example, x$^2$ is incorrect; this should be formatted as $x^2$ (or $\mathrm{x}^2$ if the romanized font is desired).
%
% Do not include text that is not math in the math environment. For example, CO2 should be written as CO\textsubscript{2} instead of CO$_2$.
%
% Please add line breaks to long display equations when possible in order to fit size of the column.
%
% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
%
% When adding superscript or subscripts outside of brackets/braces, please group using {}.  For example, change "[U(D,E,\gamma)]^2" to "{[U(D,E,\gamma)]}^2".
%
% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% Please contact latex@plos.org with any questions.
%
% % % % % % % % % % % % % % % % % % % % % % % %

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}

% amsmath and amssymb packages, useful for mathematical formulas and symbols
\usepackage{amsmath,amssymb}

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

% Use Unicode characters when possible
\usepackage[utf8x]{inputenc}

% textcomp package and marvosym package for additional characters
\usepackage{textcomp,marvosym}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref,hyperref}

% line numbers
\usepackage[right]{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% color can be used to apply background shading to table cells only
\usepackage[table]{xcolor}

% array package and thick rules for tables
\usepackage{array}

% create "+" rule type for thick vertical lines
\newcolumntype{+}{!{\vrule width 2pt}}

% create \thickcline for thick horizontal lines of variable length
\newlength\savedwidth
\newcommand\thickcline[1]{%
  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
  \cline{#1}%
  \noalign{\vskip\arrayrulewidth}%
  \noalign{\global\arrayrulewidth\savedwidth}%
}

% \thickhline command for thick horizontal lines that span the table
\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
\hline
\noalign{\global\arrayrulewidth\savedwidth}}

% Remove comment for double spacing
%\usepackage{setspace}
%\doublespacing

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in
\textheight 8.75in

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
\renewcommand{\figurename}{Fig}

% Use the PLoS provided BiBTeX style
\bibliographystyle{plos2015}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
\setlength{\headheight}{27.023pt}
\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\sf PLOS}

%% Include all macros below

\newcommand{\lorem}{{\bf LOREM}}
\newcommand{\ipsum}{{\bf IPSUM}}

%% END MACROS SECTION

\usepackage{booktabs}
\usepackage{pdflscape}

\begin{document}
\vspace*{0.2in}

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline
{ Reinforcement Learning Produces Dominant Strategies for the Iterated Prisoner's Dilemma }
% Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Marc Harper\textsuperscript{1\Yinyang},
Vincent Knight\textsuperscript{2*\Yinyang},
Martin Jones\textsuperscript{3\ddag},
Georgios Koutsovoulos\textsuperscript{4\ddag},
Nikoleta E. Glynatsi\textsuperscript{2\ddag},
Owen Campbell\textsuperscript{3\ddag},
\\
\bigskip
\textbf{1} Affiliation Google Inc., Mountain View, CA, USA
\\
\textbf{2} Affiliation Cardiff University, School of Mathematics, UK
\\
\textbf{3} Not affiliated
\\
\textbf{4} INRA, Université Côte d‘Azur, CNRS, ISA, France
\\
\bigskip

% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
%
% Remove or comment out the author notes below if they aren't used.
%
% Primary Equal Contribution Note
\Yinyang These authors contributed equally to this work.

% Additional Equal Contribution Note
% Also use this double-dagger symbol for special authorship notes, such as senior authorship.
\ddag These authors also contributed equally to this work.

% Use the asterisk to denote corresponding authorship and provide email address in note below.
* knightva@cardiff.ac.uk

\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
    We present tournament results and several powerful strategies for the Iterated
    Prisoner's Dilemma created using reinforcement learning techniques
    (evolutionary and particle swarm algorithms). These strategies are
    trained to perform well against a corpus of over 170 distinct
    opponents, including many well-known and classic strategies. All
    the trained strategies win standard tournaments against the total collection
    of other opponents. The trained strategies and one particular human made
    designed strategy are the top performers in noisy tournaments also.
\linenumbers

\section*{Introduction}

@@

Since Robert Axelrod's seminal tournament \cite{Axelrod1980a}, a number of
IPD\@ tournaments have been undertaken and are summarised in
Table~\ref{tbl:tournaments}. Further to the work described in
\cite{knight2016open} a regular set of standard, noisy~\cite{Bendor1991} and
probabilistic ending~\cite{Axelrod1980b} tournaments are carried out as more
strategies are added to the Axelrod library.
Details and results are available here:
\url{http://axelrod-tournament.readthedocs.io}. This work presents a detailed
analysis of a tournament with 176 strategies.

\begin{table}[!hbtp]
    \caption{\bf An overview of a selection of published tournaments. Not all
             tournaments were `standard' round robins; for more details
             see the indicated references.}\label{tbl:tournaments}
    \begin{center}
        \begin{tabular}{ccccc}
            \toprule
            Year     & Reference                  & Number of Strategies & Type     & Source Code\\
            \midrule
            1979     & \cite{Axelrod1980a}        & 13                   & Standard & Not immediately available\\
            1979     & \cite{Axelrod1980b}        & 64                   & Standard & Available in FORTRAN\\
            1991     & \cite{Bendor1991}          & 13                   & Noisy    & Not immediately available\\
            2002     & \cite{Stephens2002}        & 16                   & Wildlife & Not applicable\\
            2005     & \cite{kendall2007iterated} & 223                  & Varied   & Not available \\
            2012     & \cite{Stewart2012}         & 13                   & Standard & Not fully available \\
            2016     & \cite{knight2016open}       & 129                  & Standard & Fully available \\
            \bottomrule
        \end{tabular}
    \end{center}
\end{table}

@@

\section*{Materials and methods}
\subsection*{The Strategy Archetypes}

The Axelrod library now contains many parametrised strategies trained using
machine learning
methods. Most are deterministic, use many rounds of memory, and perform
extremely well in tournaments as will be discussed in the results Section.
Training will be discussed in a later section.
These strategies can encode a variety
of other strategies, including classic strategies like Tit For Tat
\cite{Axelrod1980},
handshake strategies, and grudging strategies, that always defect after
an opponent defection.

\subsubsection*{LookerUp}\label{sec:lookerup}

The LookerUp strategy is based on a lookup table and encodes a set of
deterministic responses based on the opponent's first $n_1$ moves, the
opponent's last $m_1$ moves, and the players last $m_2$ moves. If $n_1 > 0$ then
the player has infinite memory depth, otherwise it has depth $\max(m_1, m_2)$.
This is illustrated diagrammatically in Fig~\ref{fig:lookerup}.

\begin{figure}[!hbtp]
    \centering
    \includegraphics[width=\textwidth]{./assets/lookerup.pdf}
    \caption{\bf Diagrammatic representation of the Looker up Archetype.}
    \label{fig:lookerup}
\end{figure}

@@

\subsubsection*{Gambler}\label{sec:gambler}

Gambler is a stochastic variant of LookerUp. Instead of deterministically
encoded moves the lookup table emits probabilities which are
used to choose cooperation or defection.
This is illustrated diagrammatically in Fig~\ref{fig:gambler}.

\begin{figure}[!hbtp]
    \centering
    \includegraphics[width=\textwidth]{./assets/gambler.pdf}
    \caption{\bf Diagrammatic representation of the Gambler Archetype.}
    \label{fig:gambler}
\end{figure}

@@

\subsubsection*{ANN: Single Hidden Layer Artificial Neural Network}\label{sec:ann}

@@

\begin{itemize}
    \item Opponent's first move is C
    \item Opponent's first move is D
    \item Opponent's second move is C
    \item Opponent's second move is D
    \item Player's previous move is C
    \item Player's previous move is D
    \item Player's second previous move is C
    \item Player's second previous move is D
    \item Opponent's previous move is C
    \item Opponent's previous move is D
    \item Opponent's second previous move is C
    \item Opponent's second previous move is D
    \item Total opponent cooperations
    \item Total opponent defections
    \item Total player cooperations
    \item Total player defections
    \item Round number
\end{itemize}

These are then input into a feed forward neural network with one layer and
user-supplied width.  This is illustrated diagrammatically in
Fig~\ref{fig:ann}.

\begin{figure}[!hbtp]
    \centering
    \includegraphics[width=\textwidth]{./assets/ann.pdf}
    \caption{\bf Diagrammatic representation of the ANN Archetype.}
    \label{fig:ann}
\end{figure}

@@

\subsubsection*{Finite State Machines}\label{sec:fsm}

Strategies based on finite state machines are deterministic and computationally efficient.
In each round of play the strategy selects an action based on the current state
and the opponent's last action, transitioning to a new state for the next round.
This is illustrated diagrammatically in Fig~\ref{fig:fsm}.

\begin{figure}[!hbtp]
    \centering
    \includegraphics[width=\textwidth]{./assets/fsm.pdf}
    \caption{\bf Diagrammatic representation of the Finite State Machine Archetype.}
    \label{fig:fsm}
\end{figure}

Training this strategy corresponds to finding mappings of states and histories
to an action and a state. Fig~\ref{fig:fsm16} and Fig~\ref{fig:fsm16noise} show
two of the trained
finite state machines. The layout of state nodes is kept the same between
Fig~\ref{fig:fsm16} and Fig~\ref{fig:fsm16noise} to highlight the effect of
different training environments. Note also that two of the 16 states are not
used, this is also an outcome of the training process.

\begin{figure}[!hbtp]
    \centering
        \includegraphics[width=.5\textwidth]{./assets/FSM16.pdf}
        \caption{\bf Evolved\_FSM\_16: trained to maximise score in a standard
        tournament}
        \label{fig:fsm16}
\end{figure}

\begin{figure}
        \centering
        \includegraphics[width=.5\textwidth]{./assets/FSM16noise.pdf}
        \caption{Evolved\_FSM\_16\_Noise\_05: trained to maximise score in a
        noisy tournament}
        \label{fig:fsm16noise}
\end{figure}

\subsubsection*{Hidden Markov Models}\label{sec:hmm}

A variant of finite state machine strategies are called hidden Markov models
(HMMs). Like the strategies based on finite state machines, these strategies
also encode an internal state. However, they use probabilistic transitions based on the
prior round of play to other states and cooperate or defect with various
probabilities at each state. This is
shown diagrammatically in Fig~\ref{fig:hmm}. Training this strategy
corresponds to finding mappings of states and histories to probabilities of
cooperating as well as probabilities of the next internal state.

\begin{figure}[!hbtp]
    \centering
    \includegraphics[width=\textwidth]{./assets/hmm.pdf}
    \caption{\bf Diagrammatic representation of the Hidden Markov Model Archetype.}
    \label{fig:hmm}
\end{figure}

\subsubsection*{Meta Strategies}

@@

% Results and Discussion can be combined.
\section*{Results}

@@

A total of \input{./assets/number_of_players.tex}are included, of which
\input{./assets/number_of_stochastic_players.tex}are stochastic.
A standard tournament with 200
turns and a tournament with 5\% noise is
discussed. Due to the inherent stochasticity of these IPD tournaments, these
tournaments were repeated \input{./assets/standard_number_of_repetitions.tex}
times. This allows for a detailed and confident analysis of the performance of
strategies. To illustrate the results considered,
Fig~\ref{fig:tit_for_tat_scores} shows the distribution of the mean score per
turn of Tit For Tat over all the repetitions. Similarly,
Fig~\ref{fig:tit_for_tat_ranks} shows the ranks of of Tit For Tat for each
repetition. (We note that it never wins a tournament). Finally
Fig~\ref{fig:tit_for_tat_wins} shows the number of opponents beaten in any given
tournament: Tit For Tat does not win any match. (This is due to the fact that it
will either draw with mutual cooperation or defect second).

\begin{figure}[!hbtp]
        \centering
        \includegraphics[width=\textwidth]{./assets/standard_tft_scores.pdf}
        \caption{\bf Scores for Tit for Tat over
        \protect\input{./assets/standard_number_of_repetitions.tex}tournaments.}
        \label{fig:tit_for_tat_scores}
\end{figure}

\begin{figure}[!hbtp]
        \centering
        \includegraphics[width=\textwidth]{./assets/standard_tft_ranks.pdf}
        \caption{\bf Ranks for Tit for Tat over
        \protect\input{./assets/standard_number_of_repetitions.tex}tournaments.}
        \label{fig:tit_for_tat_ranks}
\end{figure}

\begin{figure}[!hbtp]
        \centering
        \includegraphics[width=\textwidth]{./assets/standard_tft_wins.pdf}
        \caption{\bf Wins for Tit for Tat over
        \protect\input{./assets/standard_number_of_repetitions.tex}tournaments.}
        \label{fig:tit_for_tat_wins}
\end{figure}

@@

\subsection*{Standard Tournament}\label{sec:standard}

@@

\begin{table}[!hbtp]
        \centering
        \footnotesize
        \input{./assets/standard_top_15_scores_summary.tex}
        \caption{\bf Standard Tournament: Mean score per turn of top 15 strategies
            (ranked by median over
        \protect\input{./assets/standard_number_of_repetitions.tex}tournaments).
        The leaderboard is dominated by the trained strategies (indicated by a
        $^{*}$).}
        \label{tbl:standard_score}
\end{table}

For completeness, violin plots showing the distribution of the scores of each
strategy (again ranked by median score) are shown in
Fig~\ref{fig:standard_boxplot}.

\begin{landscape}
    \begin{figure}[!hbtp]
        \centering
        \includegraphics[width=.8\paperwidth]{./assets/standard_scores_boxplots.pdf}
        \caption{\bf Standard Tournament: Mean score per turn (ranked by median
        over
        \protect\input{./assets/standard_number_of_repetitions.tex}tournaments).}
        \label{fig:standard_boxplot}
    \end{figure}
\end{landscape}

Pairwise payoff results are given as a heatmap (Fig~\ref{fig:standard_heatmap})
which shows that many strategies achieve mutual cooperation (obtaining a score
of 3). The top performing
strategies never defect first yet are able to exploit weaker strategies that
attempt to defect.

\begin{figure}[!hbtp]
    \centering
    \includegraphics[width=\textwidth]{./assets/standard_scores_heatmap.pdf}
    \caption{\bf Standard Tournament: Mean score per turn of row players against
    column players (ranked by median over
        \protect\input{./assets/standard_number_of_repetitions.tex}tournaments).}
    \label{fig:standard_heatmap}
\end{figure}

The strategies that win the most matches
(Table~\ref{tbl:standard_wins_top_winners}) are Defector~\cite{Axelrod1984} and Aggravater~\cite{axelrodproject}, followed
by handshaking and zero determinant strategies~\cite{Press2012}.
This includes two handshaking
strategies that were the result of training to maximize Moran process fixation
(TF1 and TF2). No strategies were trained specifically to win matches. None of
the top scoring strategies appear in the top 15 list of strategies ranked by
match wins. This can be seen in Fig~\ref{fig:standard_winplot} where the
distribution of the number of wins of each strategy is shown.

\begin{table}[!hbtp]
    \centering
        \input{./assets/standard_top_15_winners_wins_summary.tex}
        \caption{\bf Standard Tournament: Number of wins per tournament
        of top 15 strategies (ranked by median wins over
        \protect\input{./assets/standard_number_of_repetitions.tex}tournaments).}
        \label{tbl:standard_wins_top_winners}
\end{table}

\begin{landscape}
    \begin{figure}[!hbtp]
        \centering
        \includegraphics[width=.8\paperwidth]{./assets/standard_wins_boxplots.pdf}
        \caption{\bf Standard Tournament: number of wins per tournament (ranked by
        median over
        \protect\input{./assets/standard_number_of_repetitions.tex}tournaments).}
        \label{fig:standard_winplot}
    \end{figure}
\end{landscape}

@@

\begin{table}[!hbtp]
    \caption{\bf Standard Tournament: Number of wins per tournament
        of top 15 strategies (ranked by median score over
        \protect\input{./assets/standard_number_of_repetitions.tex}tournaments).}
        \label{tbl:standard_wins}
    \centering
    \footnotesize
        \input{./assets/standard_top_15_wins_summary.tex}
\end{table}

Finally, Table~\ref{tbl:standard_ranks} and
Fig~\ref{fig:standard_ranks_boxplot} show the ranks (based on median score)
of each strategy over the repeated tournaments. Whilst there is some
stochasticity, the top three strategies almost always rank in the top three. For
example, the worst that the Evolved Lookerup 2 2 2 ranks in any tournament
is 8th.

\begin{table}[!hbtp]
    \centering
        \caption{\bf Standard Tournament: Rank in each tournament
        of top 15 strategies (ranked by median over
        \protect\input{./assets/standard_number_of_repetitions.tex}tournaments).}
    \footnotesize
        \input{./assets/standard_top_15_wins_summary.tex}
        \input{./assets/standard_top_15_ranks_summary.tex}
        \label{tbl:standard_ranks}
\end{table}

\begin{landscape}
    \begin{figure}[!hbtp]
        \centering
        \includegraphics[width=.8\paperwidth]{./assets/standard_ranks_boxplots.pdf}
        \caption{\bf Standard Tournament: rank in each tournament (ranked by
        median over
        \protect\input{./assets/standard_number_of_repetitions.tex}tournaments).}
        \label{fig:standard_ranks_boxplot}
    \end{figure}
\end{landscape}
